---
rweqtitle: "Practical Machine Learning Project"
author: "Rong Cai"
date: "November 25, 2016"
output: html_document
---

##Practical Machine Learning Project

##Intruduction:

This project will use data set from accelerometers on the belt, forearm, arm, and dumbell of 6 participant to build model on the training data set and predict outcome on the test data set. I will do some pre-process on the data sets. Specially, remove some variables which are constant or missing in most situation. Then, I will divide training set into a new training set and a validation data set. I will build  two different models and validate the models on the validation data set. finally, I will pick a better model applying the model to the test data set and predict the outcome "classe".      

##Loading and Processing the data

```{r, results='hide', warning=FALSE, message=FALSE}
##Load required packages
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(gbm)
library(survival)
library(plyr)
```
```{r}
##Load data from web to my working directory
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile="./pm1-training.csv")
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile="./pm1-testing.csv")
##Read data to R
training <- read.csv("pm1-training.csv")
testing <- read.csv("pm1-testing.csv")
```

There are 19622 observations and 160 variables in training data set. There are 20 observationa and 160 variables in the testing data set. By viewing training data set, I find out there are some variables with large number of mising values and also some variables will not have any impact on our prediction project such as x, user_name, time variables, and the almost constant variables. Nest step, I will remove these variables.

```{r}
##Remove variables near zero variance to reduce variables from 160 to 100
nzv <- nearZeroVar(training)
training <- training[, -nzv]
##Remove variables with a lot of NA to reduce variables from 100 to 59
lotsNA <- sapply(training, function(x) mean(is.na(x))) > 0.90
training <- training[, lotsNA==F]
##Remove no impact variables such as ID, name...,to reduce variables from 59 to 54
training <- training[, -(1:5)]
##Do the same thing on the testong data set
nzvt <- nearZeroVar(testing)
testing <- testing[, -nzvt]
lotsNAt <- sapply(testing, function(x) mean(is.na(x))) > 0.90
testing <- testing[, lotsNAt==F]
testing <- testing[, -(1:5)]
```

Now, lets divide training data set into two data sets. one of them to use for validattion

```{r}
set.seed(123)
inTrain <- createDataPartition(y=training$classe, p=0.7, list=F)
training <- training[inTrain, ]
validation <- training[-inTrain, ]
dim(training)
dim(testing)
```

##Building Model

```{r results='hide'}
##Using three different method to build model on training set
fitControl <- trainControl(method="cv", number=3, verboseIter=F)
##Decision trees method (rpart)
modFit1 <- train(classe ~ ., data=training, method="rpart", trControl=fitControl)
##Stochastic gradient boosting trees (gbm)
modFit2 <- train(classe ~ ., data=training, method="gbm", trControl=fitControl)
##Random forest decision trees (rf)
modFit3 <- train(classe ~. , data=training, method="rf", trControl=fitControl)
```
```{r}
fancyRpartPlot(modFit1$finalModel)
modFit2
modFit3$finalModel
```

```{r}
##Prediction on validation data set
validpred1 <- predict(modFit1, validation)
validpred2 <- predict(modFit2, validation)
validpred3 <- predict(modFit3, validation)
confusionMatrix(validpred1, validation$classe)
confusionMatrix(validpred2, validation$classe)
confusionMatrix(validpred3, validation$classe)
```

From my three methods of prediction, The ramdom forest method have high accruracy. I will apply this model to test data set.   

##Generate Files to submit 

```{r}
##predict on test data set
testpred <- predict(modFit3, newdata=testing)
testpred

# create function to write predictions to files
pml_write_files <- function(x) {
    n <- length(x)
    for(i in 1:n) {
        filename <- paste0("problem_id_", i, ".txt")
        write.table(x[i], file=filename, quote=F, row.names=F, col.names=F)
    }
}

# create prediction files to submit
pml_write_files(as.character(testpred))


